\chapter{Symmetry Groups}

\section{Symmetry elements. Pole figures.}

{\bf Comment on ``It is clear that this set forms a group, the {\it
    symmetry group} of the body.''} (p.\ 32). The identity is a
symmetry. Any transformation has an inverse (move all the points back
to their original position from the transformed configuration, which
is identical to the original by definition of a symmetry
transformation) and repeated application of any two symmetry
transformation leaves the body in the same state so the product
transformation is a symmetry transformation. Thus the set forms a
group.

{\bf Point groups} (p.\ 33) comprise of only reflections in a plane and
rotations about a fixed axis and do not include translations. This
implies that all rotations and reflections leave at least one point
fixed. For example, rotations about non-intersecting axes would result
in a translation. As another example, reflections through
non-intersecting planes would also result in a translation.

{\bf Comments on ``Thus for odd $n$, if the body has the symmetry
  $S_n$, it also has the symmetries $\sigma_h$ and $C_n$ as
  independent symmetry elements.''} (p.\ 33) $S_n$ being a symmetry
means that operation by $S_n$ returns the object to its original
configuration. Thus $(S_n)^n$ is a symmetry. In addition, if $n$ is
odd then $(S_n)^n$ is $\sigma_h$. So $\sigma_h$ is a symmetry. If
$S_n$ and $\sigma_h$ are symmetries, then $S_n\sigma_h$ is a symmetry.
But $S_n\sigma_h$ = $C_n\sigma_h\sigma_h$ = $C_n$. So $C_n$ is a
symmetry. Take as an example a three-fold
arrangement above the reflection plane and one the same distance below
the plane where one is rotated by 60 degrees from the other.  This
does not have symmetry $S_3$. It does have symmetry $S_6$. It also has
symmetry $C_3$. It does not have symmetry $\sigma_h$.  Now take an
arrangement like the previous, but without the 60 degree
rotataion. This does have symmetry under $S_3$, it does have symmetry
under $\sigma_h$, and it does have symmetry under $C_3$. But note that
$n$ is odd ($n = 3$).

{\bf Comment on ``All three elements $I$, $\sigma_h$, and $C_2$
  commute.} (p.\ 37) $\sigma_h$ and $C_2$ commute (as noted on the
previous page). Remember that $I = S_2$. Since $S_2 = \sigma_h C_2$ by
definition, $I$ commutes with $\sigma_h$ and $C_2$.

{\bf Comments on ``If any two of these elements belong to the symmetry
  group, then so does the third.''} (p.\ 37) Eqs.\ (2.6) and (2.6a)
show that each of the elements can be expressed as a combination of
the other two. Also all three commute with one another. So if two are
in the group, so must the third since we can produce the third as the
product of the two.

{\bf Comment on Fig.\ 2-6.} (p.\ 37) The construction does not depend
on $P^\prime$ lying on the bisection of the angle $\phi$, as one might
infer from the drawing. In other words, point $P$ could be anywhere on
the plane, including in the interior of the angle $\phi$.

{\bf Explanation of ``...if the $X$- and $Y$-axes are 2-fold rotation
  axes, then so it the $Z$-axis.''} (p.\ 38). $n$-fold rotation axis
means that the object is symmetric under rotation by $2\pi/n$ about
that axis (defined on p.\ 33).  A rotation of $\pi$ about $X$ leaves
the object in coincidence with itself. A following rotation of $\pi$
about $Y$ also leaves the object in coincidence with itself. But the
result of these two operations is a rotation of $\pi$ about $Z$ (since
$X$ and $Y$ subtend an angle of $\pi/2$). Therefore, $Z$ is a 2-fold
rotation axis.

\section{Equivalent axes and planes. Two-sided axes.}

{\bf Explanation of ``...if the line $OP$ is an $n$-fold axis, then we
  must also have $n$-fold axes along $OP^\prime...$''} (p.\ 38) If OP
is a $n$-fold rotation axis then if a $2\pi/n$ rotation about $OP^\prime$
does not bring the object into coincidence with itself, that means
that the distribution of points with respect to $OP^\prime$ is
different than that with respect to $OP$. That contradicts the
assumption that the vertical axis is a 3-fold symmetry axis.

{\bf ...all belong to the same class...} (p.\ 38) A $2\pi/n$ rotation about
$OP^\prime$ can be obtained by a $2\pi/3$ rotation bringing $OP^\prime$ into
coincidence with $OP$, followed by a rotation of $2\pi/n$ about $OP$,
followed by the reverse rotation of $2\pi/3$ back to $OP^\prime$. So the
rotation about $OP$ by $2\pi/n$ is conjugate to that about $OP^\prime$. In a
similar manner all $n$-fold rotations about $OP$, $OP^\prime$, and
$OP^{\prime\prime}$ are conjugate to one another. They thus form a
conjugate class.

{\bf Question: What is an example of an object that has multiple equivalent axes and rotation-reflection symmetry?} (p.\ 38)
A molecule with atoms at (2,0,1), (0,2,1), (-2,0,1), (0,-2,1), (2,0,-1), (0,2,-1), (-2,0,-1), and (0,-2,-1) has four equivalent axes, the positive $x$-axis, the negative $x$-axis, the positive $y$-axis, and the negative $y$-axis. It has reflection symmetry about the $x$-$y$ plane, the $x$-$z$ plane, and the $y$-$z$ plane.

{\bf Comment on discussion of Figs.\ 2-8 and 2-9} (p.38) The rotation
about $A^\prime A$ is conjugate to that about $B^\prime B$ because we
can rotate the axis $A^\prime A$ to $B^\prime B$, perform the rotation
about $B^\prime B$, then do the reverse rotation of axis from
$B^\prime B$ to $A^\prime A$ and we recover the original rotation
about $A^\prime A$. By the same argument the $A^\prime A$ rotation is
conjugate the the rotation about $BB^\prime$.

{\bf Comment on two-sided axis definition}. (p.\ 39, 40) A two-sided
axis has a rotation and its transformed rotation differing by the
handedness of the rotation, but not the magnitude.

\section{Groups whose elements are pure rotations: uniaxial groups, dihedral groups.}

{\bf Explanation of ``Each element forms a class by itself''} (p.\ 41)
If any element is transformed by any other element, the result is the
original element. This is nearly the definition of a single element class.

{\bf Pole Figures} (p.\ 41) These are explained on p.\ 34.
